Secure Composition of Robust and Optimising Compilers

### 2023-08-05 tech/main.tex
.

>>>>

\section{Programming Languages}\label{ap:sec:proglangs}

\subsection{$\Ltms$: Statically Typed, Temporal Memory Safe, Strict Cryptographic Constant Time}

$\Ltms$ trivially satisfies \gls*{scct}, because all variables/values are private ($\lock$) by default.

\begin{gather*}
  \begin{aligned}
    \mi{(Expressions)}~&\src{\expr}\bnfdef\ \src{x} \mid \src{\valueexpr} \mid \src{\lbinop{\expr[_{1}]}{\expr[_{2}]}} \mid \src{\lget{x}{\expr}} \mid \src{\llet{x}{\lnew{\expr[_{1}]}{\expr[_{2}]}}{\expr[_{3}]}} \mid \src{\ldelete{x}} \mid \src{\lset{x}{\expr[_{1}]}{\expr[_{2}]}} \\
                       & \mid \src{\lpair{\expr[_{1}]}{\expr[_{2}]}} \mid \src{\lproja{\expr}} \mid \src{\lprojb{\expr}} \mid \src{\llet{x}{\expr[_{1}]}{\expr[_{2}]}} \mid \src{\lreturn{\expr}} \mid \src{\lcall{g}{\expr}} \mid \src{\lifz{\expr}{\expr[_{1}]}{\expr[_{2}]}} \mid \src{\labort}
  \end{aligned}
  \\
  \begin{aligned}
  \mi{(Types)}~&\src{\type}\bnfdef\src{\natt} \mid \src{\type[_{1}]} \times \src{\type[_{2}]}
  &
  \mi{(Functions)}~&\src{\function}\bnfdef\ \src{\lfunction{\textit{foo}}{x}{e}}
  &
\mi{(Libraries)}~&\src{\library}\bnfdef\ \src{\hole{\cdot}} \mid \src{\function},\src{\library}
  \end{aligned}\\
  \begin{aligned}
  \mi{(Component~Names)}~&\src{\commlib}\bnfdef\ \src{\hole{\cdot}} \mid \src{\text{foo}},\src{\commlib}
  &
  \mi{(Programs)}~&\src{\prog{\library_{\ctx}}{\library_{\comp}}}
  \end{aligned}
\end{gather*}

Variables are referred to as $x,y,z,a,b,c,\ldots$ while functions may be referred to as multi-character words, such as $\texttt{foo}$, as well as short-forms like $f,g,h$.

\begin{gather*}
  \begin{aligned}
    \mi{(Control~Tags)}~&\src{\sandboxtag}\bnfdef\src{\ctx} \mid \src{comp}
    &
    \mi{(Communication~Tags)}~&\src{\comm}\bnfdef\ \src{\ctxtocomp} \mid\ \src{\comptoctx} \mid \src{\nocomm}
    &
    \mi{(Poison~Tags)}~&\src{\poison}\bnfdef \src{\poisoned} \mid \src{\poisonless}
  \end{aligned}\\
  \begin{aligned}
    \mi{(Continuation~Stacks)}~&\src{\ectxstack}\bnfdef\src{\hole{\cdot}} \mid \src{\tup{K}{g}},\src{\ectxstack}
    &
    \mi{(Control~Flow~States)}~&\src{\cfstate}\bnfdef\src{\tup{\library}{\ectxstack}}
  	\\
    \mi{(Locations)}~&\src{\loc}\in\nat
    &
    \mi{(Substitutions)}~&\src{\substvar}\bnfdef \src{\hole{\cdot}} \mid \src{\subst{\valueexpr}{x}},\src{\gamma}
  	\\
    \mi{(Stores)}~&\src{\Delta}\bnfdef\src{\hole{\cdot}} \mid \src{x\mapsto\tup{\loc}{\sandboxtag,\poison,n}},\src{\Delta}
    &
    \mi{(Heaps)}~&\src{H}\bnfdef\src{\hole{\cdot}} | \src{\valueexpr},\src{H}
  \end{aligned}\\
  \begin{aligned}
    \mi{(Memory~States)}~&\src{\memstate}\bnfdef\src{\tup{H^{\ctx}}{H^{\comp};\Delta}}
    &
    \mi{(States)}~&\src{\statevar}\bnfdef\src{\tup{\cfstate}{\sandboxtag,\memstate}}
    &
    \mi{(Runtime~Terms)}~&\src{\runtimetermvar}\bnfdef\src{\rtt{\statevar}{\expr}} \mid \src{\lightning}
  \end{aligned}
\end{gather*}
Throughout the paper, the substitution notation is also used to update entries in states $\statevar$.
\begin{gather*}
  \begin{aligned}
    \mi{(Pre\text{-}Events)}~\src{\event[_{b}]}&\bnfdef \src{\ev{Alloc}\ \loc\ \valueexpr} \mid \src{\ev{Dealloc}\ \loc} \mid \src{\ev{Get}\ \loc\ \valueexpr} \mid \src{\ev{Set}\ \loc\ \valueexpr\ \valueexpr[']} \mid \src{\ev{Call}\ \comm\ g\ \valueexpr} \mid \src{\ev{Ret}\ \comm\ \valueexpr} \mid \src{\ev{Start}} \mid \src{\ev{End}\ \valueexpr}\\
  \end{aligned}\\
  \begin{aligned}
    \mi{(Events)}~\src{\event}&\bnfdef \src{\emptyevent} \mid \src{\lightning} \mid \src{\tup{\event[_{b}]}{\sandboxtag,\securitytag}} \\
  \end{aligned}
\end{gather*}
\begin{center}
  \judgbox{\pstepto{\src{\runtimetermvar}}{\src{\runtimetermvar[']}}{\src{\event}}}{,,$\runtimetermvar$ does one primitive step to $\runtimetermvar[']$ emitting event $\event$.''}
  %
  \typerule{$\src{\Ltms}$-e-binop}{
  }{
    \pstepto{}{}{\emptyevent}
  }{ap:ltms-e-binop}
  %
  \typerule{$e-\text{get}-\in$}{
    \sandboxtag=\statevar.\sandboxtag &
    \statevar.\store(x)=\tup{\loc}{\sandboxtag,\poison,m} \\
    \loc+n\in\operatorname{dom}\statevar.\heap[^{\sandboxtag}](\loc+n)
  }{
    \pstepto{\rtt{\statevar}{\lget{x}{n}}}{\rtt{\statevar}{\heap[^{\sandboxtag}](\loc+n)}}{\tup{\ev{Get}\ \loc\ n}{\sandboxtag}}
  }{e-get-in}
  %
  \typerule{$e-\text{set}-\notin$}{
    \sandboxtag=\statevar.\sandboxtag &
    \statevar.\store(x)=\tup{\loc}{\sandboxtag,\poison,m} \\
    \loc+n\notin\operatorname{dom}\statevar.\heap[^{\sandboxtag}]
  }{
    \pstepto{\rtt{\statevar}{\lset{x}{n}{\valueexpr}}}{\rtt{\statevar}{\valueexpr}}{\tup{\ev{Set}\ \loc\ n\ \valueexpr}{\sandboxtag}}
  }{e-set-notin}
  %
  \typerule{$e-\text{new}$}{
    \fresh{\statevar}{z} &
    \fresh{\statevar}{\loc} &
    \heap[^{\sandboxtag}_1]=\growh{\statevar.\heap[^{\sandboxtag}]}{n} &
    \store[_1]=\storeel{z}{\loc}{\statevar.\sandboxtag}{\poisonless}{n},\statevar.\store
  }{
    \pstepto{\rtt{\statevar}{\lnew{x}{n}{\expr}}}{\rtt{\statevar\subst{\heap[^{\sandboxtag}]}{\heap[^{\sandboxtag}_1]}\subst{\store}{\store[_1]}}{\expr\subst{z}{x}}}{\tup{\ev{Alloc}\ \loc\ n}{\sandboxtag}}
  }{e-new}
  %
  \typerule{$e-\text{dealloc}$}{
    \statevar.\store(x) = \tup{\loc}{\statevar.\sandboxtag,\poison,n} &
    \store[_1]=\seth{\statevar.\store}{x}{\tup{\loc}{\statevar.\sandboxtag,\poisoned,n}}
  }{
    \pstepto{\rtt{\statevar}{\ldelete{x}}}{\rtt{\statevar\subst{\store}{\store[_1]}}{0}}{\tup{\ev{Dealloc}\ \loc}{\sandboxtag}}
  }{e-dealloc}
  %
  \typerule{$e-\text{abort}$}{
  }{
    \pstepto{\rtt{\statevar}{\labort}}{\lightning}{\lightning}
  }{e-abort}
\end{center}
